Busch, Khakhalin. Intrinsic temporal tuning

### 2019-06-05 reviews and responses.md

> Page 9 line 9: “cells exposed to stronger stimuli preferred shorter synaptic inputs in dynamic clamp…”
Here, (where “preferred” is not in quotes) does “preferred” mean to fire more action potentials?

As described above, we now don't use quotation marks for any basic words, so "preferred" is used without quotation marks, and consistently means "responded preferably", or "produced stronger responses".

> Page 10 (last paragraph): Were action potential thresholds measured? Were there any differences in
this parameter across groups?

The paragraph in question read: "We ran a stepwise generalized linear model selection analysis (R package stepAIC,
Venables and Ripley 2013) to explain the intrinsic tuning of cells recorded in all
experimental groups through these eight variables. We found that the average spikiness
(after compensation for position within the tectum; see Methods) was best described by a
combination of sodium peak current (INa) and membrane resistance (Rm) variables, but
these variables explained only 8% and 2% of cell-to-cell variance respectively
F(1,130)=11.3, and F(1,130)=4.5; Figure 3E)." etc.

We measured action potential thresholds, but did not report them, as we know from earlier studies (Ciarleglio Khakhalin 2015) that AP threshold measurements in tectal neurons are highly sensitive to the current that is used to evoke the spikes, due to rapid Na channels inactivation. At the analysis phase, we briefly tried to use AP thresholds from current clamp experiments to estimate the prominence of Na channels inactivation in individual tectal neurons (Azouz 2000; Fontaine 2014), but these attempts did not lead to conclusive results, and as they are also not based on established analysis methods, they are not reported in the paper. Neither minimal threshold for generating an action potential, nor the slope of AP threshold with increase in holding current, different across groups (ANOVA p=0.7 and 0.3 respectively).

### 2019-06-07 main.tex
Text updates + diff to submission

\section*{Abstract}
% J Neurophysiol: up to 250 words
% Currently: ?

For a biological neural network to be functional, its neurons need to be connected to each other with synapses of appropriate strength, and each neuron needs to appropriately respond to its synaptic inputs. This second aspect of network tuning is maintained by intrinsic plasticity; yet it is often considered secondary to changes in connectivity, and mostly limited to adjustments of overall excitability of each neuron. Here we argue that individual neurons can also be tuned to inputs of different temporal dynamics, and that they can routinely adjust this tuning to match the statistics of their synaptic activation. Using the dynamic clamp technique, we show that in the tectum of Xenopus tadpoles, neurons become selective for faster inputs when animals are exposed to fast visual stimuli, but remain responsive to longer inputs in animals exposed to slower, looming or multisensory stimulation. We also report a homeostatic co-tuning between synaptic and intrinsic temporal properties of individual tectal cells. These results expand our understanding of plasticity in the brain, and suggest that there may exist a whole extra dimension of network tuning that was so far overlooked.

%Old version: Homeostatic intrinsic plasticity is often described as an adjustment of neuronal excitability to maintain stable spiking output. Here we report that intrinsic plasticity in the tectum of Xenopus tadpoles also supports temporal tuning, wherein neurons independently adjust spiking responses to fast and slow patterns of synaptic activation. Using the dynamic clamp technique, and five different types of visual, acoustic, and multisensory conditioning, we show that in tadpoles exposed to light flashes, tectal neurons became selective for fast synaptic inputs, while neurons exposed to looming and multisensory stimuli remained responsive to longer inputs. We also report a homeostatic co-tuning between synaptic and intrinsic temporal properties in tectal cells, as neurons that naturally received fast synaptic inputs tended to be most responsive to long-lasting synaptic conductances, and the other way around. These results expand our understanding of plasticity in the brain, and inform future work on the mechanisms of sensorimotor transformation.

% Acoustic stimuli had different effects when delivered alone, or when combined with visual stimuli.

@@

It is often assumed, in fields as diverse as connectomics and machine learning, that the main difference between functional and dysfunctional neural networks lies in their connectivity \citep{takemura2014,hildebrand2017,bassett2017network,reimann2017}. However, biological neurons also demonstrate complex and multifaceted intrinsic tuning  \citep{oleary2013,evans2015}, with cells within a single network having different activation thresholds \citep{kole2012}, bursting behaviors \citep{popovic2011}, and inactivation profiles \citep{bianchi2012}. Dysregulation of intrinsic plasticity affects the dynamics of networks  \citep{tien2018}, and can lead to a loss of function \citep{marcelin2009}. And yet, with the possible exception of oscillatory networks \citep{marder2011,picton2018control}, the exact ways in which intrinsic plasticity contributes to network tuning, remain unclear. We know that neurons adjust their spikiness to match the levels of synaptic activation they experience \citep{aizenman2003,titley2017}, but we also know that intrinsic properties affect neuronal tuning to different temporal patterns of activation \citep{azouz2000threshold,branco2010dendritic,fontaine2014threshold,jarvis2018morphology,ohtsuki2018bursting,zbili2019axonnav}. That begs the question: do neurons dynamically adjust this type of tuning in practice?

In this paper, we investigate whether exposure to sensory stimuli with different temporal dynamics would change intrinsic temporal tuning of sensory neurons in the optic tectum of \textit{Xenopus} tadpoles. Tadpole tectum is an ideal model for exploring this question: it is a highly malleable distributed network of about 10\textsuperscript{4} neurons \citep{pratt2013}, involved in stimulus discrimination and sensorimotor transformations \citep{dong2009,khakhalin2014}. In development, tectal neurons acquire diverse phenotypes, that are then shaped by sensory experiences \citep{xu2011,ciarleglio2015}. Circuits in the tectum can learn and reproduce temporal patterns to which they were exposed \citep{pratt2008}: a property that could in principle be achieved through synaptic changes alone \citep{lukovsevivcius2009}, but which may be easier to achieve through intrinsic temporal tuning \citep{narayanan2008,beatty2014}. Finally, tectal neurons exhibit strong Na channel inactivation, which seems to play a role in collision detection \citep{jang2016}, can support temporal tuning \citep{clay2012optimal,fontaine2014threshold,zbili2019axonnav}, and is a known target for plasticity mechanisms \citep{bianchi2012}.

@@

In our previously published large-scale census of tectal cells \citep{ciarleglio2015}, we observed no signs of intrinsic temporal tuning (figures 2, 4), and no interaction between intrinsic and synaptic phenotypes. We argue however, that the current clamp protocols used in earlier studies were not adequate for the task, so in this paper we rely on a dynamic clamp technique. The main benefit of the dynamic clamp, compared to commonly used current step injections, is that it allows a more realistic simulation of neuronal responses to synaptic conductances \citep{prinz2004}. In dynamic clamp, the current injected into the cell is adjusted in real-time, based on a predefined formula that depends on time and cell membrane potential. This means that with dynamic clamp recordings we can excite a neuron in a controlled manner, but still allow both its voltage and transmembrane currents to change, preserving feedback interactions between active currents and transient voltage-gated channels inactivation \citep{ma1996dynamic,zbili2019axonnav}, which is an important mechanism for temporal tuning \citep{branco2010dendritic,platkiewicz2011}. As a second methodological innovation, instead of relying on one type of sensory stimulation \citep{ciarleglio2015}, here we used five different stimulation protocols, and compared to each other their effects on intrinsic tuning.

@@

After conditioning, we excised the brain, performed whole cell patch clamp on tectal neurons (Figure 1A), and counted spikes produced in response to dynamically simulated synaptic conductances (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (\citealt{xu2011,khakhalin2014,ciarleglio2015}; see Methods). Based on prior research \citep{aizenman2003,ciarleglio2015}, we expected neurons exposed to instantaneous checkerboard inversions (flashes) to become more excitable, but surprisingly, they were on average less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, n cells=28, 29). (Here and below, F-values are reported for a multivariate fixed effects analysis of variance with selected interactions, where cell id is used as a repeated measures factor; see Methods for a detailed description). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25). This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses \citep{khakhalin2014,khakhalin2019graph}, yet their effect on neuronal excitability was weaker, seemingly in contradiction with the principle of homeostatic plasticity \citep{pratt2007,turrigiano2007}.

We then mapped the amplitude tuning of neurons (their amplitude transfer function, or gain), by looking at how an increase in transmembrane conductance translated into increased spike output. Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve (less of a difference between columns of Figure 1D, or between curves in each individual panel of Figure 1E), as they did not increase their spiking as fast in response to larger conductances (F(1,677)=15.9, p=8e-5; and F(1,629)=8.3, p=0.004 for flash- and looming respectively, compared to control). %The flattening of response curves seemed slightly more pronounced for flashes than for looming, but this difference was not significant (F(1,641)=3.5; p=0.06).

@@

We then examined whether different types of sensory activation would differently reshape temporal intrinsic tuning in tectal neurons. As changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to shorter sensory stimuli (instantaneous flashes; \citealt{khakhalin2014}) would selectively suppress spiking in response to faster synaptic inputs. As an alternative to this hypothesis, however, one could argue that exposure to shorter stimuli could make neurons better adjusted to working with fast synaptic currents \citep{stemmler1999information}, similar to changes described for synaptic plasticity \citep{aizenman2007} and recurrent activity in the tectum \citep{pratt2007, shen2011}.

We found that compared to control, neurons exposed to flashes and looming stimuli had different response curves when tested across input conductances of different lengths (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs (500 and 1000 ms) typically evoked stronger spiking than shorter inputs (100 and 200 ms), but neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike output for longer conductance injections (Figure 1E). In other words, while control neurons preferred longer inputs, stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, followed two different kinds of logic. The overall excitability was homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as "adaptive", as neurons exposed to shorter stimuli (flashes) became relatively \textit{more} responsive to shorter stimuli, and less responsive to longer stimuli, making them equipped to process faster, synchronous patterns of activation \citep{stemmler1999information, fontaine2014threshold}.

@@

To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds \citep{james2015,truszkowski2017}, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds (group “Sound”) did not lead to significant changes in either average number of spikes (0.8$\pm$0.7 spikes; F(1,689)=1.7; p=0.2; n=28, 30), amplitude transfer function (F(1,689)=2.0; p=0.2), or temporal tuning curve (F(1,689)=2.0; p=0.2). This may suggest that acoustic stimuli did not activate tectal circuits strongly enough during conditioning, despite being more behaviorally salient (at the onset of stimulation, acoustic clicks evoked startle responses in about 50-80\% of cases, compared to 5-10\% for checkerboard inversions \citep{james2015,truszkowski2017}). This was not necessarily surprising, as mechanosensory and visual inputs have different cellular targets in the tectum \citep{pratt2009trigeminal, felch2016, truszkowski2017}, potentially leading to different recruitment of tectal excitatory and inhibitory circuits. Acoustic stimuli may also be inherently weaker than visual stimuli in triggering plasticity effects in tectal neurons, as they arrive at different compartments within the dendritic tree \citep{hiramoto2009, deeg2009}, which may shape their influence on neuronal plasticity \citep{richards2019dendritic}.

We then combined visual and acoustic stimuli in two different ways and looked at the effects of multisensory stimulation on the intrinsic properties of tectal neurons. For some animals, we synchronized the instantaneous checkerboard inversions (flashes) with sound clicks (dubbed “Sync”), while for others, we staggered visual and acoustic stimuli by half a period (500 ms; dubbed “Async”). We found (Fig 1D, E) that multisensory stimulation suppressed excitability of tectal neurons less strongly than visual stimulation alone (0.6$\pm$0.4 spikes, F(1,689)=11.2, p=8e-4; and 0.7$\pm$0.6 spikes, F(1,665)=41.7, p=2e-10, for Sync and Async respectively, compared to Flash, across all testing conditions). Moreover, both temporal and amplitude tuning curves in multisensory groups were less flat than in the Flash group, with a stronger effect in the Async group (for amplitude tuning: F(1,689)=1.3, p=0.3, and F(1,665)=2.4, p=0.02 in Sync and Async groups respectively; for temporal tuning: F(1,689)=9.3, p=0.002, and F(1,665)=22.8, p=2e-6 respectively). This suggests that while on their own sound clicks had little effect on tectal excitability, when added to visual flashes, they negated some of the retuning effects that visual stimulation would have had.

@@

A visual comparison of neuronal tuning in different experimental groups (Fig 2C) shows that acoustic stimulation had opposite effects when provided on its own (without visual stimulation), compared to when it was added to visual flashes. Compared to control neurons, cells exposed to sound clicks seemed to have lower amplitude (Cohen’s d=$-$0.29) and temporal (d=$-$0.27) tuning coefficients (Hotelling t-squared test p=0.2). When sounds were added to flashes, however, both tuning coefficients increased (d=0.65 and 0.26 for amplitude and temporal tuning respectively, "Async" compared to “Flash”), making neurons more similar to control neurons (Hotelling test p=0.04). Thus, acoustic stimulation tended to tune the network in the same direction as visual stimulation when delivered alone (Figure 2D), but it negated the effects of visual stimulation the two were combined. This effect is likely to be either a consequence of strong inhibitory recruitment during multisensory integration in the tectum \citep{liu2016,hamodi2016}, or a sign of highly non-linear interactions between acoustic and visual inputs within the dendritic trees of individual tectal neurons \citep{deeg2011,felch2016,truszkowski2017}.

Somewhat similarly, looming stimuli seemed to have weaker effects on intrinsic properties, compared to flashes, both in terms of amplitude (d=0.44) and temporal tuning (d=0.22), despite them eliciting stronger responses in-vivo \citep{khakhalin2014,khakhalin2019graph}. This suggests that intrinsic retuning depends on stimuli synchronicity, rather than simply the total number of spikes generated by the network, as more synchronous stimuli (Flash, Sync) had stronger effects on tuning than comparable asynchronous stimuli (Async, Looming).

% Here we may reference Bollmann 2009 back, to show that dendritic processing probably does happen here.

@@

In contrast, \textit{within} experimental groups, cells that preferred shorter synaptic inputs in dynamic clamp tended to receive longer actual synaptic currents, and vice versa (F(1,145)=4.9, p=0.03). The deviations of neuronal properties from their respective group averages are shown in Figure 3F. Cells that received shorter synaptic currents, compared to other cells, tended to be more responsive to longer synaptic inputs (adjusted r=$-$0.19, p=0.02, n=151). This suggests that during normal brain development, individual neurons tended to tune their intrinsic properties away from the typical statistics of their inputs, enhancing responses to unusual patterns of synaptic activation. It also means that the correlation of intrinsic and synaptic properties had opposite signs between-groups (Figure 3E; positive) and within-groups (Figure 3F; negative), which is a textbook case of a so-called "Simpson's paradox" in data analysis.

@@

We can therefore conclude that our set of eight cellular parameters can much better predict spiking during current step injections (61\% of variance), compared to more physiologically realistic dynamic clamp experiments (13\% of variance). This suggests the existence of intrinsic properties that strongly affect spiking in dynamic clamp experiments, but are inaccessible through simple voltage step protocols, such as the amount of voltage-gated channels inactivation (\citealt{zbili2019axonnav}; see Discussion). Our hypothesis is also indirectly supported by the observation that of only one out of eight cellular parameterswas significantly (p$<$0.05) different across treatment groups (NaV: F(5,175)=3.7, p=0.003), and that the number of spikes detected in current clamp experiments did not differ across experimental groups (F(5,165)=0.8, p=0.6).

@@

In all analyses presented above, we adjusted cell properties for rostro-caudal and medio-lateral position of each cell within the tectum, as in tadpoles both intrinsic \citep{hamodi2014} and synaptic properties \citep{wu1996,khakhalin2012} differ between older (rostro-medial) and younger (caudal, lateral) parts of the developing tectum. Most cell properties we measured in this study correlated (p$<$0.05 after adjustment for treatment group differences) with either medial or rostral position within the tectum (medial: membrane capacitance r=$-$0.17, membrane resistance r=0.25, sodium current activation potential r=$-$0.30, stable potassium current r=$-$0.16, early synaptic amplitude r=$-$0.22, synaptic current duration r=0.29; rostral: peak sodium current r=0.18, late synaptic amplitude r=0.11; n between 168 and 183). Curiously, neither of the three measures of intrinsic tuning (average spikiness, temporal tuning, and amplitude tuning) correlated with position (p$>$0.1, n=168). This seems to suggest that while low-level properties of tectal cells depended on their developmental age, their spiking phenotypes were age-independent. It means that different cells achieved similar spiking behaviors through different combinations of underlying parameters, relying on the principle of “parameter degeneracy” \citep{prinz2004degeneracy,drion2015}.

@@

In this study we show that different sensory stimuli retuned neurons in the optic tectum of \textit{Xenopus} tadpoles in different ways, thereby inducing changes in both their temporal tuning and amplitude transfer functions. This addresses our first question about the functional scope of intrinsic plasticity in the optic tectum, and shows that it goes well beyond simple adjustments of neuronal excitability.

As our technical resources were rather limited, in this study we don't explicitly address the mechanisms of this newly discovered intrinsic temporal tuning. We can, however, offer two working hypotheses that may explain these results. As tadpole tectal cells don't express "true" resonance currents, such as h-currents \citep{ciarleglio2015}, most temporal tuning effects we observed seem to be due to differences in ionic current inactivation in different cells \citep{azouz2000threshold, fontaine2014threshold}. One obvious way of tuning ionic channel inactivation would be for every cell to modulate sodium or transient potassium currents via channel phosphorylation, or by changing the expression of different channel variants, to shift their inactivation dynamics \citep{frank2003nachannels,goldwyn2018a_current}. To test this hypothesis, one would need to induce changes in temporal tuning similarly to how we did in this paper, pharmacologically isolate different ionic currents, and directly measure their inactivation dynamics \citep{zbili2019axonnav}.

>>>>

\section*{Abstract}
% J Neurophysiol: up to 250 words
% Currently: 186

For a biological neural network to be functional, its neurons need to be connected with synapses of appropriate strength, and each neuron needs to appropriately respond to its synaptic inputs. This second aspect of network tuning is maintained by intrinsic plasticity; yet it is often considered secondary to changes in connectivity, and mostly limited to adjustments of overall excitability of each neuron. Here we argue that even non-oscillatory neurons can be tuned to inputs of different temporal dynamics, and that they can routinely adjust this tuning to match the statistics of their synaptic activation. Using the dynamic clamp technique, we show that in the tectum of Xenopus tadpoles, neurons become selective for faster inputs when animals are exposed to fast visual stimuli, but remain responsive to longer inputs in animals exposed to slower, looming or multisensory stimulation. We also report a homeostatic co-tuning between synaptic and intrinsic temporal properties of individual tectal cells. These results expand our understanding of intrinsic plasticity in the brain, and suggest that there may exist an additional dimension of network tuning that has been so far overlooked.

@@

It is often assumed, in fields as diverse as connectomics and machine learning, that the main difference between functional and dysfunctional neural networks lies in their connectivity \citep{takemura2014, hildebrand2017, bassett2017network, reimann2017}. However, biological neurons also demonstrate complex and multifaceted intrinsic tuning  \citep{oleary2013,evans2015}, where cells within a single network having different activation thresholds \citep{kole2012}, bursting behaviors \citep{popovic2011}, and inactivation profiles \citep{bianchi2012}. Dysregulation of intrinsic plasticity affects the network dynamics  \citep{tien2018}, and can lead to a loss of function \citep{marcelin2009}. And yet, with the possible exception of oscillatory networks \citep{marder2011,picton2018control}, the exact ways in which intrinsic plasticity contributes to network tuning, remain unclear. We know that neurons adjust their spikiness to match the levels of synaptic activation they experience \citep{aizenman2003,titley2017}, but we also know that intrinsic properties can affect more subtle neuronal tuning to different temporal patterns of activation \citep{azouz2000threshold, branco2010dendritic, fontaine2014threshold, jarvis2018morphology, ohtsuki2018bursting, zbili2019axonnav}. This begs the question: do neurons use this type of tuning in practice, dynamically adjusting it to the temporal dynamics of their inputs?

In this paper, we investigate whether exposure to sensory stimuli with different temporal dynamics would change intrinsic temporal tuning of sensory neurons in the optic tectum of \textit{Xenopus} tadpoles. The tadpole tectum is an ideal model for exploring this question: it is a highly malleable distributed network of about 10\textsuperscript{4} neurons \citep{pratt2013}, involved in stimulus discrimination and sensorimotor transformations \citep{dong2009,khakhalin2014}. In development, tectal neurons acquire diverse phenotypes, that are then shaped by sensory experiences \citep{xu2011,ciarleglio2015}. Circuits in the tectum can learn and reproduce temporal patterns to which they were exposed \citep{pratt2008}: a property that could in principle be achieved through synaptic changes alone \citep{lukovsevivcius2009}, but which may be easier to achieve through intrinsic temporal tuning \citep{narayanan2008,beatty2014}. Finally, tectal neurons exhibit strong Na channel inactivation, which seems to play a role in collision detection \citep{jang2016}, can support temporal tuning \citep{clay2012optimal,fontaine2014threshold,zbili2019axonnav}, and is a known target for plasticity mechanisms \citep{bianchi2012}.

@@

In our previously published large-scale census of tectal cells \citep{ciarleglio2015}, we observed no signs of intrinsic temporal tuning (figures 2, 4), and no interaction between intrinsic and synaptic phenotypes. We argue however, that the current clamp protocols used in earlier studies were not adequate for the task, so in this paper we rely on a dynamic clamp technique. The main benefit of the dynamic clamp, compared to commonly used current step injections, is that it allows a more realistic simulation of neuronal responses to synaptic conductances \citep{prinz2004}. In dynamic clamp, the current injected into the cell is adjusted in real-time, based on a predefined formula that depends on time and cell membrane potential. This means that with dynamic clamp recordings we can excite a neuron in a controlled manner, but still allow both its voltage and transmembrane currents to change, preserving feedback interactions between active currents and transient voltage-gated channel inactivation \citep{ma1996dynamic,zbili2019axonnav}, which is an important mechanism for temporal tuning \citep{branco2010dendritic,platkiewicz2011}. As a second methodological innovation, instead of relying on one type of sensory stimulation \citep{ciarleglio2015}, here we used five different stimulation protocols, and compared their individual effects on intrinsic tuning.

@@

After conditioning, we excised the brain, performed whole cell patch clamp on tectal neurons (Figure 1A), and counted spikes produced in response to dynamically simulated synaptic conductances (Figure 1B). We used conductances of 4 different durations (100, 200, 500, and 1000 ms), and 3 different amplitudes (peak conductances of 0.2, 0.5, and 1.0 nS), matching the range of synaptic currents observed in tectal circuits \textit{in vivo} (\citealt{xu2011,khakhalin2014,ciarleglio2015}; see Methods). Based on prior research \citep{aizenman2003,ciarleglio2015}, we expected neurons exposed to instantaneous checkerboard inversions (flashes) to become more excitable, but surprisingly, they were on average less spiky, and generated 0.4$\pm$0.4 spikes, across all types of dynamic clamp inputs, compared to 0.9$\pm$1.0 spikes in control (Figure 1D, left column; F(1,677)=30.4, p=5e-8, n cells=28, 29). (Here and below, F-values are reported for a multivariate fixed effects analysis of variance with selected interactions, where cell ID is used as a repeated measures factor; see Methods for a detailed description). Neurons exposed to looming transitions also spiked less than control neurons (0.6$\pm$0.4 spikes; F(1,629)=9.6; p=0.002; n=28, 25), but more than those exposed to instantaneous "flashes" (F(1,641)=3.7; p=4e-6; n=29, 25). This was likewise unexpected, as compared to flashes, looming stimuli are known to elicit stronger tectal responses \citep{khakhalin2014,khakhalin2019graph}, yet their effect on neuronal excitability was weaker, seemingly in contradiction with the principle of homeostatic plasticity \citep{pratt2007,turrigiano2007}.

We then mapped the amplitude tuning of neurons (their amplitude transfer function, or gain), by looking at how an increase in transmembrane conductance translated into increased spike output. Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve (less of a difference between columns of Figure 1D, or between curves in each individual panel of Figure 1E), as they did not increase their spiking as fast in response to larger conductances (F(1,677)=15.9, p=8e-5; and F(1,629)=8.3, p=0.004 for flash- and looming respectively, compared to control). The flattening of response curves seemed slightly more pronounced for flashes than for looming (F(1,641)=3.5; p=0.06).

@@

We then examined whether different types of sensory activation would differently reshape intrinsic temporal tuning in tectal neurons. We knew from earlier studies that fast tectal responses can be evoked with instantaneous flashes, while looming stimuli cause slower, and more prolonged tectal activation \citep{khakhalin2014, khakhalin2019graph}. As overall changes in intrinsic properties seemed homeostatic (increased activation led to reduced spiking), by the same logic, one could expect that exposure to flashes would selectively suppress spiking in response to fast synaptic inputs. On the other hand, one could argue that exposure to shorter stimuli could make neurons better adjusted to working with fast synaptic currents \citep{stemmler1999information}, similar to changes described for synaptic plasticity \citep{aizenman2007} and recurrent activity in the tectum \citep{pratt2007, shen2011}.

We found that compared to control, neurons exposed to flashes and looming stimuli had different response curves when tested across input conductances of different lengths (F(1, 677)=25.1; p=7e-7; and F(1,629)=12.0; p=6e-4 respectively). In control neurons, longer inputs (500 and 1000 ms) typically evoked stronger spiking than shorter inputs (100 and 200 ms), but neurons from stimulated animals had flatter tuning curves, with a plateau, or even a decrease in spike output for longer conductance injections (Figure 1E). In other words, while control neurons preferred longer inputs (responded to them with more spikes), stimulated neurons developed a preference for shorter synaptic inputs, and this change was more pronounced in neurons exposed to flashes than in those exposed to looming stimuli (F(1,641)=7.5; p=0.006). This suggests that the change in overall intrinsic excitability, and the change in temporal tuning, followed two different kinds of logic. The overall excitability was homeostatic, as neurons became less excitable in response to stronger stimulation. The temporal retuning however can be better described as "adaptive", as neurons exposed to shorter stimuli (flashes) became relatively \textit{more} responsive to shorter stimuli, and less responsive to longer stimuli, making them equipped to process faster, synchronous patterns of activation \citep{stemmler1999information, fontaine2014threshold}.

@@

To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds \citep{james2015,truszkowski2017}, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds (group “Sound”) did not lead to significant changes in either average number of spikes (0.8$\pm$0.7 spikes; F(1,689)=1.7; p=0.2; n=28, 30), amplitude transfer function (F(1,689)=2.0; p=0.2), or temporal tuning curve (F(1,689)=2.0; p=0.2). This may suggest that acoustic stimuli did not activate tectal circuits strongly enough during conditioning, despite being more behaviorally salient (at the onset of stimulation, acoustic clicks evoked startle responses in about 50-80\% of cases, compared to 5-10\% for checkerboard inversions \citep{james2015, truszkowski2017}). This was not necessarily surprising, as mechanosensory and visual inputs have different cellular targets in the tectum \citep{pratt2009trigeminal, felch2016, truszkowski2017}, potentially leading to different recruitment of tectal excitatory and inhibitory circuits. Acoustic stimuli may also be inherently weaker than visual stimuli in triggering plasticity effects in tectal neurons, as they arrive at different compartments within the dendritic tree \citep{hiramoto2009, deeg2009}, which may define their influence on neuronal plasticity \citep{richards2019dendritic}.

We then combined visual and acoustic stimuli in two different ways and looked at the effects of multisensory stimulation on the intrinsic properties of tectal neurons. For some animals, we synchronized the instantaneous checkerboard inversions (flashes) with sound clicks (dubbed “Sync”), while for others, we staggered visual and acoustic stimuli by half a period (500 ms; dubbed “Async”). We found (Fig 1D, E) that multisensory stimulation did not suppress excitability of tectal neurons as strongly as visual stimulation alone (0.6$\pm$0.4 spikes, F(1,689)=11.2, p=8e-4; and 0.7$\pm$0.6 spikes, F(1,665)=41.7, p=2e-10, for Sync and Async respectively, compared to Flash, across all testing conditions). Moreover, both temporal and amplitude tuning curves in multisensory groups were less flat than in the Flash group, which was especially noticeable for the Async group (for amplitude tuning: F(1,689)=1.3, p=0.3, and F(1,665)=2.4, p=0.02 in Sync and Async groups respectively; for temporal tuning: F(1,689)=9.3, p=0.002, and F(1,665)=22.8, p=2e-6 respectively). This suggests that while on their own sound clicks had little effect on tectal excitability, when added to visual flashes, they negated some of the retuning effects that visual stimulation would have had.

@@

A visual comparison of neuronal tuning in different experimental groups (Fig 2C) shows that acoustic stimulation had opposite effects when provided on its own (without visual stimulation), compared to when it was added to visual flashes. Compared to control neurons, cells exposed to sound clicks seemed to have lower amplitude (Cohen’s d=$-$0.29) and temporal (d=$-$0.27) tuning coefficients (Hotelling t-squared test p=0.2). When sounds were added to flashes, however, both tuning coefficients increased (d=0.65 and 0.26 for amplitude and temporal tuning respectively, "Async" compared to “Flash”), making neurons more similar to control neurons (Hotelling test p=0.04). Thus, acoustic stimulation tended to tune the network in the same direction as visual stimulation when delivered alone (Figure 2D), but it negated the effects of visual stimulation when the two were combined. This effect is likely to be either a consequence of strong inhibitory recruitment during multisensory integration in the tectum \citep{liu2016,hamodi2016}, or a sign of highly non-linear interactions between acoustic and visual inputs within the dendritic trees of individual tectal neurons \citep{deeg2011, felch2016, truszkowski2017}.

Somewhat similarly, looming stimuli seemed to have weaker effects on intrinsic properties, compared to flashes, both in terms of amplitude (d=0.44) and temporal tuning (d=0.22), despite them eliciting stronger responses in-vivo \citep{khakhalin2014, khakhalin2019graph}. This suggests that intrinsic retuning depends on stimuli synchronicity, rather than simply the total number of spikes generated by the network, as more synchronous stimuli (Flash, Sync) had stronger effects on tuning than comparable asynchronous stimuli (Async, Looming).

@@

In contrast, \textit{within} experimental groups, cells that preferred shorter synaptic inputs in dynamic clamp tended to receive longer actual synaptic currents, and vice versa (F(1,145)=4.9, p=0.03). The deviations of neuronal properties from their respective group averages are shown in Figure 3F. Cells that received shorter synaptic currents, compared to other cells, tended to be more responsive to longer synaptic inputs (adjusted r=$-$0.19, p=0.02, n=151). This suggests that during normal brain development, individual neurons tended to tune their intrinsic properties away from the typical statistics of their inputs, enhancing responses to unusual patterns of synaptic activation. This tuning to "unusual stimuli" fits into the narrative of information transfer maximization \citep{stemmler1999information, brenner2000} and network criticality \citep{rubinov2011}, wherein every element of a network tries to maximize its influence on the overall computation. It also means that the correlation of intrinsic and synaptic properties had opposite signs between-groups (Figure 3E; positive) and within-groups (Figure 3F; negative), which is a textbook case of a so-called "Simpson's paradox" in data analysis.

@@

We can therefore conclude that our set of eight cellular parameters can much better predict spiking during current step injections (61\% of variance), compared to more physiologically realistic dynamic clamp experiments (13\% of variance). This suggests the existence of intrinsic properties that strongly affect spiking in dynamic clamp experiments, but are inaccessible through simple voltage step protocols, such as the amount of voltage-gated channels inactivation (\citealt{zbili2019axonnav}; see Discussion). Our hypothesis is also indirectly supported by the observation that only one out of eight cellular parameters was significantly (p$<$0.05) different across treatment groups (NaV: F(5,175)=3.7, p=0.003), and that the number of spikes detected in current clamp experiments did not differ across experimental groups (F(5,165)=0.8, p=0.6).

@@

In all analyses presented above, we adjusted cell properties for rostro-caudal and medio-lateral position of each cell within the tectum, as in tadpoles both intrinsic \citep{hamodi2014} and synaptic properties \citep{wu1996,khakhalin2012} differ between older (rostro-medial) and younger (caudal, lateral) parts of the developing tectum. Most cell properties we measured in this study correlated (p$<$0.05 after adjustment for treatment group differences) with either medial or rostral position within the tectum (medial: membrane capacitance r=$-$0.17, membrane resistance r=0.25, sodium current activation potential r=$-$0.30, stable potassium current r=$-$0.16, early synaptic amplitude r=$-$0.22, synaptic current duration r=0.29; rostral: peak sodium current r=0.18, late synaptic amplitude r=0.11; n between 168 and 183). Curiously, neither of the three measures of intrinsic tuning (average spikiness, temporal tuning, and amplitude tuning) correlated with position (p$>$0.1, n=168). This seems to suggest that while low-level properties of tectal cells depended on their developmental age, their spiking phenotypes were age-independent. This means that different cells achieved similar spiking behaviors through different combinations of underlying parameters, relying on the principle of “parameter degeneracy” \citep{prinz2004degeneracy,drion2015}.

@@

In this study, we show that stimuli of different temporal dynamics differentially retune \textit{Xenopus} tectal neurons, changing both their temporal and amplitude transfer functions (Figure 2). This addresses the first question of this study, about the functional scope of intrinsic plasticity in the tectum, and shows that it goes well beyond simple adjustments of neuronal excitability.

In answer to our second question about whether intrinsic and synaptic temporal properties of tectal cells are in any way coordinated, we show that they are weakly co-tuned, and moreover, that this co-tuning is modified by sensory experience (Figure 3). In contrast with earlier studies that reported increased excitability after sensory stimulation \citep{aizenman2003, dong2009, ciarleglio2015}, we found that stimulation led to a suppression of spiking, and potentiation of fast monosynaptic inputs from the eye. We believe that the reason for this difference is that in earlier studies, the visual stimulation was provided with very bright, high-contrasted LEDs, which caused a suppression of synaptic inputs via a polyamine block of AMPA receptors \citep{aizenman2003}. This suppression then triggered a "second-order" homeostatic compensation \citep{turrigiano2011, tien2018}, making neurons spikier. In our current experiments, synaptic inputs were not suppressed, and so synaptic drive during sensory conditioning had to be stronger than in control \citep{khakhalin2014}, causing a decrease in intrinsic excitability.
